Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-13 paper.qmd
Resize figures

```{r}
#| fig-cap: "Assessing which of the 6 NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e). Layout d is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
#| fig-pos: H
#| out-width: 60%

@@

```{r}
#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h). Layout f is universally poor. Layouts a, c, g, h that show large separations between clusters are universally suboptimal. Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. The choice of best is between layouts b and e, that have small separations between oddly shaped clusters. Layout e is the best choice."
#| label: fig-pbmc-mse
#| fig-pos: H
#| out-width: 100%

@@

```{r}
#| echo: false
#| fig-cap: "XXX Number of bins for b is inadequate. Caption needs to start with what this figure is for XXX Model in \\gD{}, on the layout a and b, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$). Layout a shows three-well separated clusters with big separation, while layout b shows close clusters. In $9\\text{-}D$, the data shows three clusters with two of them being close and one separated from the others. This is why b1, b2, and b3 shows a connected edge between the close clusters, as proven by @fig-pbmc-mse as a reasonable layout. Viewing the fitted model in $9\\text{-}D$, helps to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters. Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-width: 80%

@@

```{r}
#| fig-cap: "Assessing which of the 5 NLDR layouts on the MNIST digit 1 data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e). All the layouts appear to be very similar. Layout c is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-mnist-mse
#| fig-pos: H
#| out-width: 100%

@@

<!--add langevitour screenshots and youtube animation link-->
```{r}
#| label: fig-mnist-tri-proj
#| fig-pos: H
#| fig-height: 10
#| fig-width: 10
#| fig-cap: "Model in \\gD{}, on the layout a and three views of the fit in projections from $10\\text{-}D$, for the MNIST digit 1 data ($n =  7877$ and $p = 10$). There is a big non-linear cluster and a small clusterlocated very close to one corner of the big cluster. There is a twisted pattern that is hardly visible in the static plot. Video of the langevitour animation are available at <need to add the link>."

>>>>

```{r}
#| fig-cap: "Assessing which of the six NLDR layouts on the two C-shaped clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e). Layout d is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-two_non_linear_diff_shaped_close_clusters-mse
#| fig-pos: H
#| out-width: 60%

@@

```{r}
#| fig-cap: "Assessing which of the 8 NLDR layouts on the PBMC3k data  (shown in @fig-NLDR-variety) is the better representation using MSE for varying binwidth ($a_1$). Colour  used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-h). Layout f is universally poor. Layouts a, c, g, h that show large separations between clusters are universally suboptimal. Layout d with little separation performs well at tiny binwidth (where most points are in their own bin) and poorly as binwidth increases. The choice of best is between layouts b and e, that have small separations between oddly shaped clusters. Layout e is the best choice."
#| label: fig-pbmc-mse
#| fig-pos: H
#| out-width: 60%

@@

```{r}
#| echo: false
#| fig-cap: "XXX Number of bins for b is inadequate. Caption needs to start with what this figure is for XXX Model in \\gD{}, on the layout a and b, and two views of the fit in projections from $9\\text{-}D$, for the PBMC3k data ($n =  2622$ and $p = 9$). Layout a shows three-well separated clusters with big separation, while layout b shows close clusters. In $9\\text{-}D$, the data shows three clusters with two of them being close and one separated from the others. This is why b1, b2, and b3 shows a connected edge between the close clusters, as proven by @fig-pbmc-mse as a reasonable layout. Viewing the fitted model in $9\\text{-}D$, helps to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters. Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-width: 60%

@@

```{r}
#| fig-cap: "Assessing which of the 5 NLDR layouts on the MNIST digit 1 data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e). All the layouts appear to be very similar. Layout c is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-mnist-mse
#| fig-pos: H
#| out-width: 60%

@@

<!--add langevitour screenshots and youtube animation link-->
```{r}
#| label: fig-mnist-tri-proj
#| fig-pos: H
#| fig-height: 10
#| fig-width: 10
#| out-width: 60%
#| fig-cap: "Model in \\gD{}, on the layout a and three views of the fit in projections from $10\\text{-}D$, for the MNIST digit 1 data ($n =  7877$ and $p = 10$). There is a big non-linear cluster and a small clusterlocated very close to one corner of the big cluster. There is a twisted pattern that is hardly visible in the static plot. Video of the langevitour animation are available at <need to add the link>."
